### 2017-06-01 01-intro.tex
Text thingy in Related

Starting from an existing implementation in Python/Numpy, we create an efficient
baseline version in C++ to serve as the baseline for the parallel versions.
This already improves performance by a factor of 1000 to 1500 compared to
Python. Then, our key contributions are:
\begin{itemize}
\item
our implementation is scaled up to run on GPUs.
We developed aggressive optimizations to solve the memory bandwidth issues for
GPUs. This implementation achieves speedup factors up to~120,
compared to the baseline.
%
Through careful analysis of the computation and data structures we show that
the algorithm's full state can be reduced by roughly 75\%. Compressing the
state significantly reduces the data intensity and allows for tackling larger
problems; % while maintaining all state in memory.

>>>>

Starting from an existing implementation in Python/Numpy, we created an efficient
baseline version in C++ to serve as the baseline for the parallel versions.
This already improves performance by a factor of 1000 to 1500 compared to
Python. Then, our key contributions are:
\begin{itemize}
\item
our implementation is scaled up to run on GPUs.
We developed aggressive optimizations to solve the memory bandwidth issues for
GPUs. This implementation achieves speedup factors up to~120,
compared to the baseline.
%
Through careful analysis of the computation and data structures we show that
the algorithm's full state can be reduced by roughly 75\%. Compressing the
state significantly reduces the data intensity and allows for tackling larger
problems; % while maintaining all state in memory.

### 2017-06-01 02-02-related.tex
Text thingy in Related

We consider related work for both parts of this paper: acceleration and scaling
of our MCMC-MMSB algorithm. Acceleration of deep learning algorithms on GPUs
is an ongoing success story. Although approximative Baysian algorithms are
not such a natural fit for GPUs, a number of projects explores this terrain.
Medlar~c.s.~\cite{journals/bioinformatics/MedlarGSBK13} use GPUs with their
MCMC approach to analyze parental linkage patterns in a biology context and
White and Porter~\cite{DBLP:journals/csda/WhiteP14} do the same to model terrorist activity.
Latent Dirichlet Allocation, another variety of Bayesian Approximation,
is used on GPUs by Yan~c.s.~\cite{DBLP:conf/nips/YanXQ09}. There is also related
work on the variants of MCMC algorithms that use the gradient to speed
up convergence. Langevin and Hamiltonian dynamics are representatives of
these varieties~\cite{Girolami_riemannmanifold}. Our algorithm uses Riemann
Manifold Langevin dynamics. Beam~c.s.~\cite{beam2014fast} use GPUs to perform
Hamiltonian descent, using Python interfaces to access the standard cuBLAS
library~\cite{cuBLAS}. They limit their optimizations for the GPU to saving
on data transfers between host and device memory. They do not apply their
framework to MMSB.

>>>>

We consider related work for both parts of this paper: acceleration and scaling
of our MCMC-MMSB algorithm. Whereas acceleration of \emph{deep learning algorithms} on GPUs
is an ongoing success story, approximative Baysian algorithms are
not such a natural fit for GPUs. Nevertheless, a number of projects explores this terrain.
Medlar~c.s.~\cite{journals/bioinformatics/MedlarGSBK13} use GPUs with their
MCMC approach to analyze parental linkage patterns in a biology context and
White and Porter~\cite{DBLP:journals/csda/WhiteP14} do the same to model terrorist activity.
Latent Dirichlet Allocation, another variety of Bayesian Approximation,
is used on GPUs by Yan~c.s.~\cite{DBLP:conf/nips/YanXQ09}. There is also related
work on the variants of MCMC algorithms that use the gradient to speed
up convergence. Langevin and Hamiltonian dynamics are representatives of
these varieties~\cite{Girolami_riemannmanifold}. Our algorithm uses Riemann
Manifold Langevin dynamics. Beam~c.s.~\cite{beam2014fast} use GPUs to perform
Hamiltonian descent, using Python interfaces to access the standard cuBLAS
library~\cite{cuBLAS}. They limit their optimizations for the GPU to saving
on data transfers between host and device memory. They do not apply their
framework to MMSB.
